Kindly Bent to Free Us

### 2019-07-04 kindly/motivation.tex
more sudoku

\cref{sig:hybarray} contains an API \lstinline/HybArray/ 
that enables using mutable and immutable modifications to
the board through affine types and borrows.
The signature differs slightly from the \lstinline/Array/
signature. For simplicity, the array elements are unrestricted, 
hence it is sufficent for the array to be affine.
The in-place mutation function \lstinline/set_mut/ with type
\lstinline/&!('a t) * int * 'a -> unit/ works on a mutable borrow  whereas the persistent
\lstinline/set/ operation has type \lstinline/&('a t) * int * 'a -> 'a t/. It
takes an immutable borrow because it only reads from the
argument array, but returns a fresh,  modified structure.
The code in \cref{ex:cow} contains a very simple implementation of
\lstinline/HybArray/ that represents hybrid arrays
as regular arrays and uses copy-on-write for persistent
modifications. The function
\lstinline/create_from : int -> (int -> 'a) -> 'a Array.t /
initializes an array from a function.
% A more elaborate implementation could use persistent arrays
% \cite{DBLP:conf/ml/ConchonF07}.

\begin{figure}[tp]
  \centering
  \begin{subfigure}{0.48\linewidth}
\begin{lstlisting}
module type HybArray = sig
  type ('a : un) t : aff

@@

The typing ensures that the mutations do not compromise the state at
the choice point, because they operate on a new state \lstinline/new_g/ created for one
particular branch of the choice.
In particular, the \lstinline/try_solution/ function can be used in a loop
precisely because the \lstinline/set/ function only requires an
unrestricted immutable borrow of \lstinline/g/, which can be used as
many times as needed.
We could even free \lstinline/new_g/ in \cref{line:free:new_g}.

@@

\begin{figure}
  \centering
  \begin{minipage}[t]{0.45\linewidth}
    \lstinputlisting[linerange=1-13]{code/sukodu.affe}
  \end{minipage}\hfill
  \begin{minipage}[t]{0.54\linewidth}
    \lstinputlisting[linerange=15-28,firstnumber=last]{code/sukodu.affe}
  \end{minipage}
  \caption{Excerpt of the Sudoku solver}
  \label{ex:sudoku}
\end{figure}

>>>>

\cref{sig:hybarray} contains an API \lstinline/HybArray/ 
that enables using mutable and immutable modifications to
the board through affine types and borrows.
The signature differs slightly from the \lstinline/Array/
signature. For simplicity, the array elements are unrestricted, but
the structure itself remains linear so as to be implemented in terms of \lstinline/Array/.
The in-place mutation function \lstinline/set_mut/ with type
\lstinline/&!('a t) * int * 'a -> unit/ works on a mutable borrow  whereas the persistent
\lstinline/set/ operation has type \lstinline/&('a t) * int * 'a -> 'a t/. It
takes an immutable borrow because it only reads from the
argument array, but returns a fresh,  modified structure.
The code in \cref{ex:cow} contains a very simple implementation of
\lstinline/HybArray/ that represents hybrid arrays
as regular arrays and uses copy-on-write for persistent
modifications. The function
\lstinline/create_from : int -> (int -> 'a) -> 'a Array.t /
initializes an array from a function.
% A more elaborate implementation could use persistent arrays
% \cite{DBLP:conf/ml/ConchonF07}.

\begin{figure}[tp]
  \centering
  \begin{subfigure}{0.48\linewidth}
\begin{lstlisting}
module type HybArray = sig
  type ('a : un) t : lin

@@

The typing ensures that the mutations do not compromise the state at
the choice point, because they operate on a new state \lstinline/new_g/ created for one
particular branch of the choice.
In particular, the \lstinline/try_solution/ function can be used in a loop
precisely because the \lstinline/set/ function only requires an
unrestricted immutable borrow of \lstinline/g/, which can be used as
many times as needed.
As \lstinline/g/ is linear we must free it in \cref{line:free:g}.

@@

\begin{figure}
  \centering
  \begin{minipage}[t]{0.45\linewidth}
    \lstinputlisting[linerange=1-13]{code/sukodu.affe}
  \end{minipage}\hfill
  \begin{minipage}[t]{0.54\linewidth}
    \lstinputlisting[linerange=15-29,firstnumber=last]{code/sukodu.affe}
  \end{minipage}
  \caption{Excerpt of the Sudoku solver}
  \label{ex:sudoku}
\end{figure}
